### 2014-06-13 response/response.tex
added new data for responses

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs.tex}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{../paper}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Fast sparse period estimation: Reviewer responses}
\author{R.~G.~McKilliam, I.~V.~L.~Clarkson and B.~G.~Quinn 
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\maketitle

\section*{Reviewer 1}\label{sec:reviewer-1}

\begin{enumerate}

\item\textbf{Comment:}
This paper proposes a new fast method for estimating the period of sparse and noisy
observations of a point process. The new idea is to use the periodogram of quantized
observations that can be computed faster than the standard periodogram by using
a chirp z-transform or a fast Fourier transform. The reviewer is not convinced that
there is enough material deserving publication in the IEEE signal processing letters.
Moreover, there are several important comments that should be addressed before a
possible publication of this letter.
\\
\textbf{Response:}
We agree with the reviewer that the results in this paper are simple.  We do not believe this to be a weakness. Despite being simple, the results are novel and useful.  The quantisated periodogram estimator described in this paper is as accurate as the most accurate estimators in the literature (the least squares and periodogram estimator).  At the same time our estimator is substantially faster to compute.  For these reasons, we firmly believe this will become the estimator of choice for periodic point processes.  For the first time, it enables highly accurate period estimation for extremely large data sets in reasonable time. This was not possible with previous estimators that require a number of operations growing quadratically with data size. %We have provided all the simulation code as supplementary material.  This will enable simple implementation for future research.

\item\textbf{Comment:}
The main reviewer concern is about the choice of the quantization parameter
$q$ that should be explained more carefully. It is not sufficient to display two sets
of curves obtained for two different values of $q$ and to choose the best one without
explanations. For instance, it is not clear whether the value of $q$ chosen in this paper
will be appropriate for other noise distributions or for other ways of generating the
integers $s_n$. This point should be addressed carefully before a possible resubmission
of the paper. The other reviewer comments are summarized below.
\\
\textbf{Response:}
We anticipate that developing a rigorous theoretical method for choosing $q$ will be extremely complicated.  The properties of many optimisation proceedures, such as the Newton-Raphson method, and Brent's method~\cite[Ch.~5]{Brent_opt_no_derivs_1973} are reasonably well understood for the purpose of maximising fixed known functions. However, very few theorectical statistical analyses for these proceedures exist when the functions to be maximised are themselves random variables, i.e., when noise is involved.  Add to this the existence of quantisation (itself typically very tricky to analyse) and it appears that a theorectical analysis of this estimator will be extremely difficult.

These theoretical difficulties do not negate the immediate practical applicability of our estimator.  We believe it would be unwise to halt dissemination of these practical results simply because a difficult theoretical problem has not yet been solved.  We agree with the reviewer that the simulations presented are limited. This is entirely due to the 4 page limit required IEEE Signal Processing Letters.  In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantisation parameter.  It is natural to question whether this choice also works in other scenarios, for example (as the reviewer suggests), for different noise distributions, and different methods of generating the integers $s_1,\dots,s_N$.  The results of such simulations are displayed in Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document. The simulations are setup as in the paper, but with the following modifications:

\begin{enumerate}
\item In Figure~\ref{plot:poisson} the noise is Gaussian distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right. 
\item In Figure~\ref{plot:geomuniform} the noise if uniformly distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right. 
\end{enumerate}

Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ is a good one. With this choice the accuracy of the quantised periodogram estimator is similar to the original periodgram estimator (computed without quantisation and FFT).  %That the choice $q=5 f_{\text{max}}$ works in a wide range of scenarios should not come as a

\begin{figure*}[tp] 
  \centering 
  \begin{tikzpicture} 
    \selectcolormodel{gray} 
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200poisson1}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200poisson1};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5poisson1};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0poisson1};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200poisson1};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200poisson1}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200poisson1};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson1};
     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
   \end{axis} 
 \begin{axis}[name=plot2,at={(10.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200poisson10}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200poisson10};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5poisson10};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0poisson10};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200poisson10};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200poisson10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200poisson10};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson10};
     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
   \end{axis} 
  \end{tikzpicture}  
  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussain noise and $N=200$ observations.  The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.}\label{plot:poisson}
\end{figure*}

\begin{figure*}[tp]  
  \centering
  \begin{tikzpicture}  
    \selectcolormodel{gray} 
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200uniformgeom1}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200uniformgeom1};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5uniformgeom1};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0uniformgeom1};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200uniformgeom1};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200uniformgeom1}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200uniformgeom1};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom1};
     \legend{Periodogram theory, Least squares theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
   \end{axis}   \begin{axis}[name=plot2,at={(4.7,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200uniformgeom10}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200uniformgeom10};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5uniformgeom10};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0uniformgeom10};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200uniformgeom10};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200uniformgeom10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200uniformgeom10};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom10};
     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
   \end{axis} 
  \end{tikzpicture}  
  \caption{Mean square period error versus noise variance $\sigma^2$ with uniform noise and $N=200$ observations.  The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.  The least squares estimator is the most accurate in this simulation. The is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.}\label{plot:geomuniform}
\end{figure*}

\item\textbf{Comment:}
p. 3, the authors mention “Sadler and Casey [24] \dots However, their estimator
requires far larger transforms and produces less accurate results than the estimator
described here (see Section VI)". Unfortunately, there is nothing in
Section VI (which is the conclusion) supporting this claim. The authors should
compare their results in term of accuracy and of computational complexity to
the results of [24]. Including the Cramer-Rao bounds for the estimation of T0
(as in [24]) would also be appreciated.
\\
\textbf{Response:}
Section VI in the paper~\cite{726812} entitled ``On periodic pulse interval analysis with outliers and missing observations'' is not the conclusion.  It is the section entitled ``Numerical Results''.  In Figures 1 to 4 in~\cite{726812} the periogram estimator implemented by Casey and Sadler is seen to be inferior to their MEA estimator.  Casey and Sadler do use an FFT to compute the periodgram, but fail to implement an optimisation proceedure.  We used the Newton-Raphson proceedure in the old manuscript, but have used Brent's method~\cite{Brent_opt_no_derivs_1973} in the new manuscript as dicussed in our response to comment~\ref{com:rev1:nr} below.  Casey and Sadler arrive at the conclusion that very large FFTs are required and that the periodgram estimator is not particularly accurate.  In the conclusion of~\cite{726812} it is stated that:

``Comparisons with the periodogram show the limitations of periodogram performance.''

Our paper clearly shows that such limitation do not exist.  The periodgram estimator, computed by quanitisation and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.  For moderately large $N$ this is the most computationally efficient proceedure currently available for this problem.  Simultaneously, it is as accurate at the most accurate estimators available, namely the least squares estimator, and the periodgram estimator computed directly without FFT.

In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial. To the authors knowledge, a Cramer-Rao bound is not known in this case.   The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.  This bound is therefore not valid for the simulations we present in Figure~\ref{plot:multipleN} in Section~\ref{sec:simulations} of the paper.

In Figure~\ref{plot:crb} in this document we perform a simulation with integers $s_1,\dots,s_N$. The integers are generated once such $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed.  This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.  Observe that for small noise variance the \emph{clairvoyant} Cramer-Rao bound is similar to the asymptotic variance of the least square and periodogram estimator predicted in~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.  As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the asymptotic variances diverge.  As expected, the behaviour of the estimators themselves follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.

%We prefer to show to plot the MSE of the estimatorsThe least squares estimator is the maximum likelihood estimator, and so, the dashed curve serves a similar type of purpose as the CRB often does in these types of simulations.  Namely, to indicate whether an estimator is performing similarly to what might be considered `optimal', i.e. close the what might theoretically be expected from the maximum likelihood estimator.

\begin{figure*}[tp]  
  \centering
  \begin{tikzpicture}  
    \selectcolormodel{gray} 
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200fixed1}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200fixed1};
      \addplot[mark=none,color=black,dotted] table {code/data/CRBN200fixed1};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5fixed1};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0fixed1};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200fixed1};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200fixed1}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200fixed1};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200fixed1};
     \legend{Periodogram theory, Least squares theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, Least squares, SLS2-new}
     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
   \end{axis}   \begin{axis}[name=plot2,at={(4.7,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200fixed10}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200fixed10};
      \addplot[mark=none,color=black,dotted] table {code/data/CRBN200fixed10};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5fixed10};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0fixed10};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200fixed10};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200fixed10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200fixed10};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200fixed10};
     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
   \end{axis} 
  \end{tikzpicture}   
  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=200$ observations.  A singled fixed set of integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.  This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.  In this case the  \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) exists and is displayed by the dotted line.  In this simulation the \emph{clairvoyant} Cramer-Rao bound is equal to the variance of the maximum likelihood (i.e. least squares) estimator in the case that the integers $s_1,\dots,s_N$ are known.}\label{plot:crb}
\end{figure*}

\item\textbf{Comment:}\label{com:rev1:nr}
p. 7, the authors indicate that they have used the Newton-Raphson (NR)
method. More details should be provided here. For instance, what is the exact
NR scheme used in this paper (there are several options)? How have the
authors selected the total number of iterations? Is the quantized periodogram
just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?
Also, do the results displayed in Fig. 2 correspond to the total time
required to compute the estimator (including the NR procedure) or just to the
initialization of the NR procedure?
\\
\textbf{Response:}
We are only aware of a single Netwon-Raphson method that proceeds by the iteration
\[
x_{n+1} = x_n - \frac{f^\prime(x_{n})}{f^{\prime\prime}(x_n)}
\]
to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.  As was stated Section~\zref{sec:simulations}: ``Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.''  This was the case both for the simulations presented in Figure~\zref{plot:multipleN} and for the computational benchmarks presented in Figure~\zref{plot:benchmark}.

To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.  This optimiation method is likely to be more familiar with readers because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.

BLERG: In Section~\zref{sec:peri-estim} the text has been modifed to read:

``An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite{Brent_opt_no_derivs_1973} 
can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.''

\item\textbf{Comment:}
p. 5, in the definition of $v_m$, $b_{m+\ell_1}$ should be replaced by $b_{m+\ell_{\text{min}}}$.
\\
\textbf{Response:}
Thankyou for spotting this error.  A similar error was fixed in equation for $I_z$ directly above this.

\item\textbf{Comment:}
p. 6, ``each of length $L+K-1$'' might be replaced by ``each of length $M = L + K - 1$"
\\
\textbf{Response:}
We have not done this.  The parameter $M$ is reserved for the length of the FFT used for the single Fourier transform method.  We feel it will lead to confusion if $M$ is also used for the length of the transforms required for the chirp-z transform.

\item\textbf{Comment:}
p. 8, $n = 2,\dots,N$ should be $n = 1, \dots, N - 1$
\\
\textbf{Response:}
This error has be fixed.  The same error in the caption of Figure~\zref{plot:multipleN} has been fixed.

\item\textbf{Comment:}
p. 8, it is not completely clear to identify the curves associated with $N = 30$ and $N = 1200$. Can the authors display 4 curves associated with $\mu= 1$, $\mu = 10$, $N = 30$ and $N = 1200$?
\\
\textbf{Response:}
We have connected the data points for the $N=1200$ case by gray lines.  We believe this make the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.

\item\textbf{Comment:}
Typos: p. 2, “anomolies" should be “anomalies", p. 7, “quanitised" should be
“quantized".
\\
\textbf{Response:}
Done.

\item\textbf{Comment:} 
Is it necessary to include 11 self citations in this letter? Removing some of
these references would allow to save some space, which could be used to better
explain some points mentioned above.
\\
\textbf{Response:}
IEEE Signal Processing Letters allows a 5th page that can contain only references.  Therefore, we cannot save space by removing references.   We have removed reference [29] from the original manuscript because the material in that paper is also covered in [27].  We feel that the all 37 references in the new version of the paper are highly relevant.  The paper would be weakened without them.

\end{enumerate}

\section*{Reviewer 2}

\begin{enumerate}

\item\textbf{Comment:}
 The identifiability claim at the beginning of sec. II is not obvious and a reference is mandatory. Indeed, referring to eq. (2), let us assume that $s_n = 2 i_n$ where $i_n$ are all non-negative integer. In this case the point process is periodic with period $2 T_0$. This referee does not believe that maximizing the periodogram over $(f_{\text{min}},f_{\text{max}})$ will identify the true frequency (or period). LS methods might have the capabilities to identify the period, but, in any case the claim of idenitifiability needs to be further supported.
\\
\textbf{Response:}

\item\textbf{Comment:}
Is the equality in the second (un-numbered) equation of page 2 correct?
\\
\textbf{Response:}
Yes.  The terms inside the sum on the left and right are conjugates of one another.  That is,
\[
\sum_{n=1}^N e^{ 2\pi j f y_n} = \sum_{n=1}^N (e^{-2\pi j f y_n})^* = \left( \sum_{n=1}^N e^{ -2\pi j f y_n} \right)^*
\]
where $^*$ denotes the complex conjugate.  So,
\[
I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2 = \babs{ \sum_{n=1}^N e^{ -2\pi j f y_n} }^2.
\]

\item\textbf{Comment:}
Newton-Rapson iteration to further improve the frequency estimation is advocated in Sec. II. Can a reference supporting this statement be added?
\\
\textbf{Response:}
BLERG

\item\textbf{Comment:}
Referring to Fig. 3, in case $N=30$, the novel method incurs in a significant loss of performance with respect to classical LS and periodogram for low-signal-to-noise ratios. In this scenario (low $N$ and low signal-to-noise-ration) relevant for real applications?
\\
\textbf{Response:}
The quantised periodgram estimator does not suffer a significant performance loss with respect to the classical LS and periodogram estimators in any scenario we have found.  Observe in Figure 3 that the solid dots ($q=5 f_{\text{max}}$) the crosses (LS), the circles (periodogram) are all close for $N=30,1200$ and $\mu=1,10$.  Be aware that the solid lines indicates the asymptotic theory ($N\to\infty$)~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est} for the periodogram and LS estimators and not the measured accuracy for finite $N$.  As $N\to\infty$, the MSE of the periogram and LS estimators will converge to the solid and dashed curves.

\end{enumerate}

{\small
\input{../paper.bbl}
}

\end{document}