### 2014-08-20 response/proof_corrections.tex
added proof corrections

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs.tex}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Fast sparse period estimation: Reviewer responses}
\author{R.~G.~McKilliam, I.~V.~L.~Clarkson and B.~G.~Quinn 
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

%\maketitle

\begin{enumerate}

\item Page 1, column 2, \emph{periodogramestimator} appears to be written as all one word.  \\ It should be \emph{periodogram~estimator}.

\item Page 2, column 1, again \emph{periodogramestimator} appears to be written as all one word.  \\ It should be \emph{periodogram~estimator}.

\item Page 2, column 1.  The equation
\[
G = \{ f_{\text{min}} + Dk \mid k = 0, \dots K-1 \}
\]
In the proof version there no space between $Dk$ and the vertical bar $\vert$ and $k$.  It should have some space as above.

\item Page 2, bottom of column 2.  The equation
\[
b_i = \#\{ n \mid \ell_n = i, n = 1,\dots, N\}
\]
In the proof version there no space between $n$ and the vertical bar $\vert$ and $\ell_n$.  It should have some space as above.

\newcommand{\ellmin}{\ell_{\text{min}}}
\newcommand{\ellmax}{\ell_{\text{max}}}

\item Page 3, top of column 1.  The equation
\[
I_z(f) = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 = \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ellmin} }^2.
\]
does not quite fit on one line in your format.  I don't personally mind this, but if it is preferable you can split this equation over two lines, such as,
\begin{align*}
I_z(f) &= \babs{ \sum_{i = \ellmin}^{\ellmax} e^{ -2\pi j f i/q} b_i }^2 \\
&= \babs{ \sum_{m = 0}^{\ellmax-\ellmin} e^{ -2\pi j f m/q} b_{m+\ellmin} }^2.
\end{align*}

\end{enumerate}

\end{document}