### 2015-04-10 poster/ITR_Poster.tex
added start of ICASSP poster

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}

\usepackage[orientation=portrait,size=a1,scale=1.1]{beamerposter}

\mode<presentation>{
    \usetheme{ITR}    
}

%%% fonts & encoding %%%
\usepackage{newcent}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}

\input{presdefs}

\definecolor{darkgreen}{rgb}{0,0.3,0}
\newcommand{\term}[1]{{\color{darkgreen}\textbf{#1}}}

%%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Noncoherent least squares estimators \\ of carrier phase and amplitude}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, \\ I. Vaughan L. Clarkson and Barry G. Quinn}
\newcommand{\email}{robby.mckilliam@unisa.edu.au}
\newcommand{\worksupportedby}{Supported under the Australian Government's Australian Space Research Program}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%\frame{

%\vspace{-1.4cm}

\begin{columns}[t] % <-- columns with top alignment
\begin{column}{.49\textwidth} % <-- start the first column

\newcommand{\calC}{{\mathcal C}}
\begin{block}{Received signal model}
Receive $L$ noisy $M$-ary phase-shifted-keyed ($M$-PSK) symbols of the form
\[
y_i = a_0 s_i + w_i, \qquad i = 1, \dots, L.
\]
\begin{itemize}
\item $a_0 = \rho_0 e^{j\theta_0}$ is the unknown carrier phase $\theta_0$ and amplitude $\rho_0$,
\item $s_1,\dots,s_L$ are transmitted $M$-PSK symbols,
\item $w_1,\dots,w_L \in \complex$ are i.i.d. circularly symmetric complex random variables representing noise.
\end{itemize}

Interested in estimating $\rho_0$ and $\theta_0$.
\end{block}

\begin{block}{Least squares estimator}
If all symbols $s_1,\dots,s_L$ are known
\[
\hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*.
\]
More interested in the practical situation where symbols are not known,
\[
\hat{a} = \arg\min_{a \in \complex} \min_{s_1, \dots, s_L} \sum_{i = 1}^L \abs{ y_i - a s_i }^2.
\]
This estimator can be computed in $O(L\log L)$ operations.
\end{block}

\begin{block}{Theorem (Almost sure convergence)}

Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
%\begin{equation}\label{eq:RiandPhii}
\vspace{-0.15cm}
\[
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
\]
%\end{equation}
and define the continuous function
\[
G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\] 
If $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
\begin{enumerate}
\item \vspace{-0.15cm} $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item \vspace{-0.2cm} $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$,
\end{enumerate}
where $\sfracpart{\cdot}$ takes its argument `modulo $\tfrac{2\pi}{M}$' into $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
%\[
%\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
%\]  
\end{block}

\begin{block}{Theorem (Asymptotic normality)}
Let $f(r,\phi)$ be the joint pdf of $R_1$ and $\Phi_1$, and let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
\]
Put $\hat{\lambda}_L = \sfracpart{\hat{\theta} - \theta_0}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$. %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then 
The distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
\[
\left( \begin{array}{cc} 
H^{-2} A & 0 \\
0 & \rho_0^2 B
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\vspace{-0.15cm}
\[
H = G(0) -  2 \sin\big(\tfrac{\pi}{M}\big) \sum_{k = 0}^{M-1} g\big(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}\big),
\]
\vspace{-0.15cm}
\[
A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1} - G^2(0). 
\]

\end{block}

\end{column}
\begin{column}{.49\textwidth} % <-- start the second column

\begin{block}{Simulations}

\begin{figure}
	\centering
		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
		\caption{Phase mean square error for $4$-PSK (QPSK)}
		\label{fig:plotphaseM4}
\end{figure}

\vspace{0.79cm}
\begin{figure}
	\centering
		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
		\label{fig:plotampM4}
\end{figure}

\end{block}

\end{column}
\end{columns}

 %}

\end{document}